---
title: "heeia_test"
author: "Nakoa Farrant"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(deSolve)
library(ggplot2)
library(schoolmath)
library(tidyverse)
library(TropFishR)
require(MASS)
require(dplyr)
library(colorspace)
library(MLmetrics)
library(lubridate)
```

```{r}
# The below values are characteristic of Mugil cephalus (striped mullet, ‘ama‘ama) as defined in the S.I. of Cheung, et al. 2013, who derived these values from data that is freely available at fishbase.org
Linf_mugil = 60 # asymptotic length for Mugil cephalus, [cm]
K_mugil = 0.27 # von Bertalanffy growth parameter, [1/year]
a_mugil = 0.032 # This a is specific to M. cephalus in Hawai'i estuaries (Peyton, et al. 2015) The a length-weight conversion parameter for overall M.cephauls on fishbase is 0.0085
b_mugil = 3.03 # This b is specific to M. cephalus in Hawai'i estuaries (Peyton, et al. 2015). For general fishbase, the length-weight conversion parameter is b = 3 for M. cephalus
Winf_mugil = a_mugil*Linf_mugil^b_mugil # asymptotic weight using length-weight conversion [grams]
# Using the Pauly 1979 empirical equation for t0 from fishbase. Seems to be log10 and not ln because other instances on the fishbase page explicityly use ln.
t0_mugil = -exp(-0.3922-0.2752*log10(Linf_mugil)-1.038*log10(K_mugil))

# For Mugil cephalus (striped mullet), natural mortality can be empirically derived using the method developed by Then, et al. 2015
M_mugil = as.numeric(M_empirical(Linf = Linf_mugil, K_l = K_mugil, method = "Then_growth")) # [1/year]
```

```{r, message=FALSE}
# USGS Stream Gauge #16275000 He'eia Stream at Haiku Valley near Kaneohe, Oahu, HI
# Data originally collected in cubic feet per second and converted to m^3 / day
# Data can be found at: https://waterdata.usgs.gov/nwis/monthly?referred_module=sw&amp;site_no=16275000&amp;por_16275000_41087=2643571,00060,41087,1914-02,2019-05&amp;format=html_table&amp;date_format=YYYY-MM-DD&amp;rdb_compression=file&amp;submitted_form=parameter_selection_list
daily_stream <- read_csv('heeia_stream_daily.csv') 

st_clean <- daily_stream %>% 
  mutate(mean = mean_va*0.028*1.157*10^5) %>% # variable to convert mean flow from ft^3/s to m^3/d
  mutate(date = as.Date(paste(month, day, sep = "-"), "%m-%d")) %>% # this attaches the 2019 year even though these are average flow rates from 1914 to 2019
  filter(date != 2019-02-29) %>%  # remove Feb 29 that's only present in leap years
  mutate(JD = 1:365) %>% # create Julian Day variable
  dplyr::select(date, JD, mean)


# Monthly solar irradiance [W/m^2] near He‘eia Fishpond based on Univ. of Hawai'i data
# Monthly solar radiation rasters provided the data at the point 157.808ºW, 21.435ºN
# Data source: solar.geography.hawaii.edu/downloads.html
daily_solar <- read_csv('solar_irrad.csv') 

```


# 6-Hour Time Step: Model Equations
Realized that running into issues with the stream and tide specifications that are determined based on the value of t and need to be adjusted based on the time step being taken
```{r}

fish_func6 <- function(t, state6, pars6) {
  with(as.list(c(state6, pars6)), {
    
    # create an integer counter variable, time is otherwise in 6 hr intervals
    step = t*365*4 
    
    # the current year is defined as the rounded up integer year value of the current time step
    yr_curr = ceiling(t) # e.g. t= 0.5 is considered year 1. t = 1.2 is considered year 2. 
    
    # define flow terms- m^3/d
    
# There are 4 tidal cycles every day that alternate between flood and ebb tides at each 6 hour time step. Additionally, roughly every 8 days there is a switch from spring tide to neap tide based on the moon phase (spring when full or new moon; neap during quarter moon phases). This results in desired shifts from spring to neap every ~32 tide shifts (8 days * 4 tides/day)
    
# Every lunar cycle (approximately a month), there are two sets of spring and two sets of neap tides which totals to 128 tidal cycles. By using the modulo operator based on 64 tidal cycles, you can get all of the spring and neap cycles and all of the flood and ebb cycles in an approximately 2 week period that repeats twice each month.

# We are initializing this model to conditions on Jan. 1, 2014. The moon phase was a new moon representing the start of a spring tide cycle. The tide was high (flood conditions) in the early morning hours of 1/1/2014, so we begin with a spring flood tide and iterate through the cycle from there.
    
    if (step %% 64 < 33){
      # start with the first 32 spring tide cycles
      if (as.integer(step) %% 2 == 0)
        # alternate between spring flood and spring ebb based on being an even or odd step
        Q_tide = sp_fl
      else
        Q_tide = sp_ebb
    }
    else{
      # shift to  32 neap tides
      if(as.integer(step) %% 2 == 0)
        # alternate between neap flood and neap ebb based on being an even or odd step
        Q_tide = np_fl
      else
        Q_tide = np_ebb
    }

    
    # Specify He'eia stream flow rates based on USGS gauge daily flow averages from 1914-2019
    # Q_st is defined below in units of m^3 / 6 hours from m^3 / day
      if(step %% 365 != 0)
        Q_st = st_clean$mean[step %% 365]/4 
      else
        Q_st = st_clean$mean[365]/4 # account for Dec 31 case when step = 365 so 365 %% 365 would equal 0
      
      # Use the number of years that have passed to determine the nitrogren loading based on the relative 
      #changes in the land use to the other land uses from the current condition. After 20 years, each of 
      # the scenarios will have reached their expected loading rate per that land use.
      if(t < 20){
        W_No = (LU_No + No_inc*yr_curr)*Q_st # loading rate of organic N into the stream
        W_Na = (LU_Na + Na_inc*yr_curr)*Q_st # loading rate of ammonium N into the stream
        W_Ni = (LU_Ni + Ni_inc*yr_curr)*Q_st # loading rate of nitrate N into the stream
      } else {
        W_No = (LU_No + No_inc*20)*Q_st # loading rate of organic N into the stream
        W_Na = (LU_Na + Na_inc*20)*Q_st # loading rate of ammonium N into the stream
        W_Ni = (LU_Ni + Ni_inc*20)*Q_st # loading rate of nitrate N into the stream
      }

      # Fish weight as a function of time based on von Bertalanffy growth parameters [g]
      W_fish = Winf_mugil*(1-exp(-K_mugil*(t-t0_mugil)))^b_mugil 
    
      resp_o2 = 10^(-0.17*log10(W_fish) + 2.49) # [mg O2 kg fish^-1 hr^-1] based on fish base equation for weight dependent relationship for respiration rate of M. cephalus
      c_resp = resp_o2*RQ*c_mw/o2_mw*10^-6 # [g C g fish^-1 6hr^-1]
      FC = c_resp*c_energy/c_mw*6 # [kJ g fish^-1 6hr^-1]  respiration energy per fish
      Na_exc = c_resp/alg_CN # [g N-NH3 g fish^-1 6hr^-1] nitrogen excretion in the form of ammonium
      #FB = *N # [g fish/m^2] Biomass of fish in 
      
      
      ## PHYTOPLANKTON GROWTH - primarily derived from Burford and Lorenzen 2002
      # growth rate: g = g_alg_max * L_light * L_N 
      # phytoplankton growth is at its maximum except when limited by light or nutrients (nitrogen)
      g_alg_max = 0.5/4 # [6 hr^-1] max phytoplankton growth rate for 6 hour period ~1.45 was used in the Lorenzen study and 1.5 is used elsewhere. 0.5 is used in the Turner NPC study
      k_chl = 14 # [m^-1 g Chl ^-1] light extinction caused by Chl-a
      k_other = 2.5 # [m^-1] light extinction caused by non Chl-a factors
      c_chl = 30 #  ratio of carbon to Chl-a in algae by mass
      #k_ext = k_chl*alg/c_chl*A_fp + k_other# [m^-1] light extinction coefficient
      k_ext = 0.0138 # [m^-1] from Lorenzen, C.J., 1972
      z = h_fp/2 # [m] average depth of He‘eia Fishpond
      
      # Monthly solar irradiance [W/m^2] near He‘eia Fishpond based on Univ. of Hawai'i data
      if(step %% 365 != 0)
        I0 = daily_solar$irradiance[step %% 365]/4 
      else
        # account for Dec 31 case when step = 365 so 365 %% 365 would equal 0
        I0 = daily_solar$irradiance[365]/4  
    
      I_sat = 60 # [W/m^2] saturating sunlight intensity for phytoplankton growth. 
      # Converted from 275 μmol photons m−2 s−1 (Gao, et al, 2007) using this unit converter: 
      # http://www.egc.com/useful_info_lighting.php
      
      # Calculate light limitation factor
      L_light = exp(1)/k_ext*(exp(-I0/I_sat*exp(-k_ext*z))) - exp(-(I0/I_sat))
      
      k_SN <- 0.015 # [g/m^3] half saturation for nitrogen. 
      
      L_N <- (Na_fp + Ni_fp)/(Na_fp + Ni_fp + k_SN) # Nitrogen limiting component of algae growth
      
      N_Chl <- 13 # N/Chl-a ratio of algae

      g_alg <- g_alg_max*L_light*L_N
      
      
      # Stream organic N [g/m^3]  
      dNo_st <- (W_No - Q_st*No_st)/V_st -k_oa*No_st + k_ao*Na_st
      # Stream ammonium [g/m^3]  
      dNa_st <- (W_Na - Q_st*Na_st)/V_st + k_oa*No_st - (k_ao+k_nit+k_av)*Na_st + k_nit_red*Ni_st
      # Stream nitrate [g/m^3]  
      dNi_st <- (W_Ni- Q_st*Ni_st)/V_st + k_nit*Na_st - k_nit_red*Ni_st
      # Fishpond water organic N [g/m^3]  
      dNo_fp <- (Q_st*No_st + Q_tide*(No_ocean-No_fp))/V_fp + (-k_oa-k_s)*No_fp+k_ao*Na_fp + k_r*No_sd
      # Fishpond water ammonium [g/m^3]     
      dNa_fp <- (Q_st*Na_st + SGD_Na + Q_tide*(Na_ocean-Na_fp))/V_fp - k_oa*No_fp - (k_ao+k_nit+k_av+k_s)*Na_fp + k_r*Na_sd + ((-g_alg + k_ra)*alg + (eg + exc)*fish)/alg_CN/h_fp*(Na_fp/(Na_fp+Ni_fp))
      # Fishpond water nitrate [g/m^3]  
      dNi_fp <- (Q_st*Ni_st + SGD_Ni + Q_tide*(Ni_ocean-Ni_fp))/V_fp + k_nit*Na_fp-k_s*Ni_fp+k_r*Ni_sd + ((-g_alg + k_ra)*alg + (eg + exc)*fish)/alg_CN/h_fp*(Ni_fp/(Na_fp+Ni_fp))
      # Sediment organic N [g/m^3]   
      dNo_sd <- k_s*No_fp - (k_oa+k_b+k_r)*No_sd + k_ao*Na_sd + k_s*alg/alg_CN/h_fp*(No_fp/(Na_fp+No_fp))
      # Sediment ammonium [g/m^3]  
      dNa_sd <- k_s*Na_fp + k_oa*No_sd - (k_ao+k_nit_sed+k_r+k_b)*Na_sd + k_nit_red*Ni_sd + k_s*alg/alg_CN/h_fp*(Na_fp/(Na_fp+No_fp))
      # Sediment nitrate [g/m^3]    
      dNi_sd <- k_s*Ni_fp+k_nit_sed*Na_sd-(k_denit+k_nit_red+k_r+ k_b)*Ni_sd
      # Algae density [g/m^2]
      
      Ing <- FA/alg_energy/alg*fish # [day^-1], ingestion rate of algae by fish based on M. cephalus anabolic requiements and energy density of algae
    
      #dalg <- (k_ag*(Na_fp+Ni_fp)/(k_SN+Na_fp+Ni_fp) -k_ra)*alg - FA*FB/alg_energy 
      dalg <- 0.001*g_alg*alg - k_s*alg - k_ra*alg - 0.0001*Ing*alg
      # works with the current ingestion rate and the g_alg*0.001 (before adding SGD)
      
    
      # using a density dependent natural mortality rate for algae k_am*alg*alg = k_am*alg^2 see Franks (2002)
      
      # How should settling algae enter the sediment? organic nitrogen or another form?
      
      # FB increases as W increases but N also decreases. If N was at 5000 and weight reached Winf_chanos, then the last term in this equation would be - 9.39 which is more than the initial alg concentration
      # Algae stocking rate in M. cephalus hatchery experiment in Makapuu, HI would stock with 500,000 cells/mL of algae (phytoplankton Nannochloropsis oculata)
      
      # Zooplankton equation
      #dZ <- h*e-h*alg*Z-k_zm*Z^2
      
      # Benthos equation- is it more appropriate for the benthos to be interacting with the water column or sediment nitrogen sources
      #dB <- k_bg*B/(1+k_bh*B)*(No_sed + Na_sed + Ni_sed) - k_bm*B^2
    
      # Fish density
      
      # Account for no fishing during closed seasons and before any of the fish are mature
      # Also account for fish density changes during spawning (lost mass for energy going to reproduciton. Also gained mass from new fish though they aren't ready to harvest)
      # M. cephalus reproductive season is September through March 
      # (https://naturalhistory2.si.edu/smsfp/irlspec/Mugil_cephal.htm) 
      # Going to use HI DLNR closed season for ‘ama‘ama (Dec 1 to March 31)
      # (https://dlnr.hawaii.gov/dar/files/2016/11/mullet_handout_estuaries.pdf)
      
      # fish are reproducing after 2 years
      if (t > 2) {
        if(step %% 366 < 91 | step %% 366 > 334) { 
          catch_rate = 0 # Closed season (Dec 1 to March 31)
          # there should be fish mass losses due to reproduction
        }
        else{
          catch_rate = k_catch # harvest after 2 years during times of year outside of the closed season
        }
      }
      else {
        catch_rate = 0 # no harvesting if there are no fish that are at least 2 years old
      }
      
      # Account for recruitment following reproduction
      if(step < 365*3 && t %% 365 != 0)
        # No recruits until 3 years after stocking when the fish born after 2y are 1year old
        # No recruits except for the start of a new year
        R = 0 
      else
        # number of recruits to the population 
        # N is already in units of fish per m^2. Thus, R has units of [# of 1 yr survivors/female] * [# of females / m^2 ] = # of 1 year survivors / m^2
        #R = recr_perf*N/2
        R=0
      
      #dfish <- FA/alg_energy/alg*fish^2 - M*fish^2 - catch_rate*fish^2 
      # dfish <- FCR*FA/alg_energy/alg*fish^2 - M*fish^2 - catch_rate*fish
      #dN <- -(M+catch_rate)*N #+ R  # [number of fish/m^2] the number of fish per m^2
      # need to introduce some kind of reproduction or other increase in fish
      # dFE = (FA - FC)**N - (M+catch_rate)*FE # [kJ/day]  
      #dfish <- FA/alg_energy/alg*fish^2*(1-fish/Fcarr) - M*fish^2 - catch_rate*fish^2 
      
      dfish <- Ing*alg*fish*(1-fish/Fcarr) - M*fish^2 - catch_rate*fish^2 
      
      f_sgr = 1.02/4 # [6 hr^-1] Wassef, et al 2002. specific growth rate suggested 2% day^-1
      #dfish <- f_sgr*fish*(1-fish/Fcarr) - M*fish^2 - catch_rate*fish^2 
      #dfish <- K_mugil*(Winf_mugil-fish) - catch*fish # somewhat of a VBGF output
      dcatch = catch_rate*fish*A_fp # if use "- catch" does it remove the catch from the previous year so it is just the current catch? This runs and need to think about the result more critically # do you need to add the new catch to the previous catch "+ catch" this doesn't seem to produce a functional result. When including this "+ catch" the equation produces weird negatie results
      
      #dcatch = catch*fish*A_fp/1000 # catch in kg (was originally in grams/m^2)
      
      # The below equations are from: Lorenzen, K. 2008 "Fish Population Regulation Beyond "Stock and   
      # Recruitment": The Role of Density-dependent Growth in the Recruited Stock
      #dLinf = Linf_mugil - g*B # biomass density dependent maximum length for VBGF
      #dL = Linf - (Linf - L)*exp(-K_mugil) # biomass density dependent length for M. cephalus with VBGF
      #dN = N*exp(-(Fc + M)) # population number
      #dFc = Finf/(1 + exp(q*(L-Lc))) # fishing mortality given by logistic function of length
      #dQ = 1/(1 + exp(p*(L-Lm))) # proportion of mature fish
      #dB = a_mugil*(L^b_mugil)*N # total biomass of the population
      #dS = a_mugil*(L^b_mugil)*N # spawner biomass of the population
    
    return(list(c(No_st = dNo_st, Na_st = dNa_st, Ni_st = dNi_st, No_fp = dNo_fp, Na_fp = dNa_fp, Ni_fp = dNi_fp, No_sd = dNo_sd, Na_sd = dNa_sd, Ni_sd = dNi_sd, alg = dalg, fish = dfish, catch = dcatch)))

})
}
```

# Model Specification
```{r}
years = 500

t <- seq(0, years, by = 1/365/4) # 6 hour time step

yr_ct <- rep(0:years, each = 365*4)

# define physical parameters of ecosystem 
h_st = 0.5 # [m] average depth of stream 
l_st = 700 # [m] length of interest for He‘eia Stream
w_st = 15 # [m] approximate width of stream

A_fp = 360000 # [m^2] area of fishpond
h_fp = 0.4 # [m] average depth of fishpond water column [m]

A_sed = 360000 # [m^2] area of sediment below fishpond water column
h_sed = 0.005 # [m] depth of sediment of interest in meters

# define volumes of model's three physical compartments [m^3]
V_st = h_st*l_st*w_st # stream
V_fp = A_fp*h_fp # fishpond water
V_sed = A_sed*h_sed # fishpond sediment

# Nitrogen loading from submarine groundwater discharge. Estimated from Dulai, et al 2016
SGD_Na <- 0 # [g/d] NH4+ -N loaded into He'eia Fishpond
SGD_Ni <- 0 # [g/d] NO3- -N loaded into He'eia Fishpond

# 670 and 74 for total ~744

# Nitrogen reaction rates
k_oa = 0.01/4 # [1/6hr] conversion of organic N to ammonium N
k_ao = 0.1/4 # [1/6hr] conversion of ammonium N to organic N 
k_nit = 0.5/4 # [1/6hr] nitrification (ammonium N -> nitrate N) in water column
k_nit_red = 0.5/4 # [1/6hr] reduction of nitrate N to ammonium N. Originally 0.1/4
k_av = 0.7/4 # [1/6hr] ammonium to ammonia (gas) volatilization rate
k_nit_sed = 0.05/4 # [1/6hr] nitrification (ammonium N -> nitrate N) rate in the sediment
k_denit = 0.01/4 # [1/6hr] denitrification rate (nitrate to N2 gas)

# Fishpond water-sediment exchange rate constants
k_r = 0.8/4 # [1/6hr] resuspension rate
k_s = 0.8/4 # [1/6hr] settling rate, used to be 0.05 /day. If going to make it this high, maybe it should just be related to phytoplankton settling rate due to their size. Burdoff and Lorenzen suggest 0.8 /day
# figures look a little different with 0.05 /day but still retain generally the same shape
k_b = 0.001/4 # [1/6hr] burial removal rate [1/6hr] 

# fish and algae rates and constants
r_NChl = 0.2 # [g N/g Chla] ratio of nitrogen to Chl-a in algae
k_ra = 0.1/4 #0.025/4, # [1/6hr] algae respiration rate
k_ag = 1.5/4 # [1/d] algae growth rate constant at 20ºC
#k_fg = 0.015/4, # rate of fish growth due to algae consumption [1/6hr]
# K for mullet is 0.27 1/yr -> 0.0007 1/day, this is lower than the current growth rate but the current growth rate also seems particularly low
k_ad = 0.09/4 # [1/6hr] rate of algae death, Jamu and Piedrahita, 0.01-0.09 1/d range
k_ac = 0.08 # expect M. cephalus to consume ~8% of it's body weight in food, De Silva & Perera 
k_fg = 0.27/365/4 # [1/6hr], seems like this is too low of a growth rate if the population is going to approach the carrying capacity
#k_fg = K_mugil/365/4, # rate of fish growth [1/6hr], K_mugil = 0.27 
# based on doubling time of 1.4-4.4 years, r_growth = ln(2)/1.4 = 0.495 /yr to ln(2)/4.4 = 0.158 /yr, prior r (whatever that means) is 0.51
M = M_mugil/365/4 # [1/6hr] rate of fish death from empirical Then formula, M_mugil = 0.346
k_catch =0.7/365/4 # could be 0.343/yr based on Mexico study.
# F_msy = 1.2 /year if harvest at 2 years using the 60cm Linf
# file:///Users/Nakoa/Downloads/2016GallardoetalfisherydeMugelcephalusJMBO-5-151.pdf
# fishbase also says you generally want Exploitation E to be about 0.5 which is F/Z = F/(F+M) so M and F should be approximately equal ideally for MSY
# From Barman, Jana, Garg, Bhatnagar, https://link.springer.com/article/10.1007/s10499-004-2479-5
# growth rate for M. cephalus in salinity 20 ppt 0.69 g/d or 0.67 g/d in 25 ppt
# SGR --specific growth rate of weight-- (% BW (in grams) per day) about 4% per day
k_birth = 0.35/365/4 # [1/6hr]

# nitrogen concentrations in ocean - units g/m^3
No_ocean = 0.001
Na_ocean = 0.04 
Ni_ocean = 0.02

# water volume flux rate (m^3/6hr)
sp_fl= 191660/4
sp_ebb = -174880/4
np_fl = 141384/4
np_ebb = -159938/4

# FISH CONSUMPTION ENERGY AND MASS PARAMETERS
# 126 kJ gross energy (GE) / kg body weight for daily maintenance but 1446.7 kJ GE / kg BW is recommended for maximum growth
FA = 1000/1000/4 # [kJ g fish^-1 6h^-1] energy consumed per gram of M. cephalus every 6 hours from FAO website says maximum growth rate is 1447 kJ/kg BW daily and 126 kJ/kg BW for maintenance
alg_energy = 17.6 # [kJ g^-1 algae] energy per gram of Nannochloropsis spp. from Rebolloso-Fuentes, et al 2001
# microplankton energy density 8 kJ g^-1 in Mediterranean (Barroeta, et al 2017). Whyte 1987 seems to further support ~8 kJ g^-1

FCR = 4.6 # feed conversion ratio inputs/outputs [alg/fish]. average 3.6-5.6 for 3% and 5% BW feeding ratios. This at 4.6 FCR could be high compared to some other studies

w_avg = 1000 # [g] average weight of M. cephalus at adult age

RQ = 0.82 # respiration coefficient from Serpa, et al, 2013 dimensionless
c_mw = 12.011 # [g] atomic weight of carbon
o2_mw = 32 # [g] molecular weight of oxygen gas
c_energy = 373 # [kJ mol C^-1]

alg_CN = 6.6 # ratio of C:N in algae based on the Refield ratio 106:16

adult_energy = 7.8 #[kJ/g Wet Mass] energy content of 23 cm adult mugil cephalus which is about a 390 g fish which is 3040 kJ per fish from Marais and Erasmus, 1977
# alternative energy content conversion for M. cephalus on Fishbase under the "Mass Conversion" tab: 5.3 kJ/g WM

# FISH EGESTION AND EXCRETION ENERGY AND MASS PARAMETERS
a_eg = 0.16 # egested fraction of consumption from Megrey, et al. 2007 (0.16 recommended by source)
a_ex = 0.1 # excreted percentage of consumed food that isn't egested from Megrey, et al. 2007 (0.1 recommended by source)
eg = a_eg*FA/alg_energy # [g alg g fish^-1 6hr^-1] egestion rate
exc = a_ex*(FA/alg_energy - eg) # [g alg g fish^-1 6hr^-1] excretion rate

# FISH DENSITY
l_fingerling = 3 # [cm] stock pond using 3cm fingerlings of M. cephalus
fingerling_mass = a_mugil*l_fingerling^b_mugil # [g] approximate mass of fingerlings at time of stocking

fingerling_energy = adult_energy*fingerling_mass # [kJ per fingerling]

Fcarr = 56 # [g/m^2] carrying capacity for fish in He‘eia Fishpond, 500 pounds per acre
max_number_dens = Fcarr/Winf_mugil # [fish/m^2] carrying capacity in number of fish based on the maximum weight of M. cephalus based on von Bertalanffy growth

## RECRUITMENT
egg_per_mass <- 750 # eggs/g of fish mass that females mature and reproduce starting 2-4 years (2 year weight ~990 g, 4yr weight ~2600g)

# M. cephalus are considered mature at 2 years
egg_per_f <- egg_per_mass*Winf_mugil*(1-exp(-K_mugil*(2-t0_mugil)))^b_mugil # egg mass per female fish at 2 years of age

# reproductive season in Hawai'i is between September and March. This State gov. website narrows that season to Dec. through Feb. https://dlnr.hawaii.gov/dar/files/2014/04/fishes_of_hawaii.pdf

# 22% survival rate after 50 days. Adjust with the
recr_perf <- egg_per_f*0.22*M_mugil # juveniles per female surviving 1 year post hatching as they are recruited tot he fishery
# need to multiply by N/2 (approximate # of females in population) at the appropriate time point to 

# NITROGEN LOADING
LU_No = 1.8 # [g/m^3] organic N concentration in wetland for current land use
LU_Na = 0.07 # [g/m^3] ammonium N concentration in wetland for current land use
LU_Ni = 0.07 # [g/m^3] nitrate N concentration in wetland for current land use

# Parameterization for other land use scenarios
# Annual increases in loading of nitrogen forms to He‘eia Stream based on the potential upstream land uses. Here we use linear increases in loading for the restored and urban scenarios such that the 
pars_control <- c(
      No_inc = 0, # [g/m^3]
      Na_inc = 0, # [g/m^3]
      Ni_inc = 0 # [g/m^3]
)

pars_restored <- c(
      No_inc = 0.22, # [g/m^3]
      Na_inc = 0.009, # [g/m^3]
      Ni_inc = 0.009 # [g/m^3]
)

pars_urban <- c(
      No_inc = 0.56, # [g/m^3]
      Na_inc = 0.023, # [g/m^3]
      Ni_inc = 0.023 # [g/m^3]
)

# there is seasonal variability in the baseline nitrogen forms present in the stream, fishpond, and sediment based on tides, timing of nitrogen applications to the lo‘i, and precipitation, among other factors.

# Information on nitrogen output from Kako‘o taro agriculture from Bremer, et al 2018: total N is 0.16 
# (+/- 0.08) g/m^3 with nitrate N 0.006 (+/- 0.014) g/m^3 and ammonia N at 0.006 (+/- 0.006) g/m^3
# these represent an order of magnitude smaller ammonia and nitrate N composition from values before

# Nitrogen in FP water column (no measures of organic N):
# summer mean:
# nitrate = 0.66 g/m^3
# ammonium = 0.02 g/m^3

# winter mean:
# nitrate = 0.69 g/m^3
# ammonium = 0.05 g/m^3

# low measured levels of nitrite (0.01 and 0.02 g/^3 in winter and summer, respectively)

# Sediment nitrogen concentrations based on the Briggs, et al paper from 2013 on coastal endmembers
# conversion factor for NH4-N: 1 mg NH4/l = 55.4 umol NH4/l
# In the first 1cm of sediment:
# ammonium (in terrogenous near egret island): 0.54 g/m^3 (~30 uM) - it's more like 25uM at most in the first 1cm
# ammonium (in carbonate sediment in more central part of pond): 0.63 g/m^3 (~35 uM)

# 0.59 g/m^3 average

# organic N: 
# waiting on organic N in FP and sediment, nitrate in sediment.


state6 <- c(
           No_st = 0.148, # [g/m^3] stream initial organic N
           Na_st = 0.006, # [g/m^3] stream initial ammonium
           Ni_st = 0.006, # [g/m^3] stream initial nitrate
           No_fp = 0.1, # [g/m^3] fishpond initial organic N, estimated from Charles Young thesis
           Na_fp = 0.03, # [g/m^3] fishpond initial ammonium, originally 0.035. The mean for the data is 0.02
           Ni_fp = 0.68, # [g/m^3] fishpond initial nitrate, originally 0.675
           No_sd = 0.05, # [g/m^3] fishpond sediment initial organic N, estimated from Young thesis
           Na_sd = 0.45, # [g/m^3] fishpond sediment initial ammonium. Sediment is a source according to Hargreaves (1998)
           Ni_sd = 0.01, # [g/m^3]  fishpond sediment initial nitrate. Sediment is a sink according to Hargreaves (1998).
           alg = 0.018, # [g/m^2] was 0.018. initial algae concentration based on average of data from Frank, et al, 2016 (0.0015 g Chl a/m^3) to g/m^2 algae. 0.0015 g Chla/m^3 * 30 carbon/Chl * 0.4
           # in 2007, Heeia FP algae level would also be about 0.018 g/m^2, essentially the same average as found in 2014-2016
           # carbon to chlorophyll ratio (w/w) is 30 (Parsons et al 1984)
           # previously suggested algae to Chl ratio was 0.02 g algae/g Chla from the literature
           # range of conversion factors [g alg/g Chla] from Chapra 0.005-0.025. Would be 0.12 g/m^2 algae if used 0.005
           #N = 5000/A_fp,
           #FB = fingerling_mass*5000,
           fish = 3.2, # [g/m^2] initial ‘ama‘ama  density (was 3.2)
           catch = 0)

# According to Mohlenkamp et al. 2019, Paepae o He'eia produced ~1.2 metric tons of moi (Pacific Threadfin) from 2006-2009, which is about 400 kg per year
# historical estimates for this fishpond max out around 15 metric tons (15,000 kg) with an asymptotic weight of 8.5 kg per striped mullet this would amount to ~1750 fully grown striped mullet.

# instead a 3kg mullet would be quite large according to Hi‘ilei which would require 5,000 fully grown mullet (and more because of natural deaths)

# According to FishBase, the common length is 50 cm (Peyton, et al 2015 corroborates this for adult M. cephalus in Hawai'i) which equates to a weight closer to 1 kg based on the length-weight conversion factor.  This would require closer to 15,000 fish to reach capacity

# hypothetically stock with 3 cm fingerlings that have a mass of ~230 grams each based on length-weight conversion. If we stock the fishpond with 5,000 fingerlings, the inital mass in the fishpond is 1.15*10^6 grams (1.1 metric tons). Initial fish density [g/m^2] 1150000 g/360000m^2 = 3.19 g/m^2 = 3.2 g/m^2

# volume = 144000m^3 => 1150000 g/144000m^3 = 7.986 g/m^3 = 8 g/m^3

# fish base indicates sexual maturity at 3-4 years for Mugil cephalus. At the end of 4 years, fish are estimated to be ~3.2 kg each, which would translate to 15 metric tons

# Biswas, et al. 2012 (https://www.sciencedirect.com/science/article/pii/S0044848612000658) suggests a stocking density of M. cephalus fry (~0.17 g in size) at 15,000 fry/hectare. He'eia fishpond is 36 ha. This is 91,800 grams of M. cephalus stocked in the fishpond, which is close to what was suggested above with slightly larger fish

# 1150000 g/144000m^3 = 7.986 g/m^3
# fish carrying capacity at 500 pounds per acre -> 56 g/m^2

```

```{r}
out_control <- ode(y = state6, times = t, func = fish_func6, parms = pars_control, method = "euler")
#out_restored <- ode(y = state6, times = t, func = fish_func6, parms = pars_restored, method = "euler")
#out_urban <- ode(y = state6, times = t, func = fish_func6, parms = pars_urban, method = "euler")
```


```{r}
plot(out_control)
# expect ammonium in the fishpond to be about 0-0.15 g/m^3
# expect nitrate in the fishpond to be about 0.5-1.5 g/m^3
```

```{r}
plot(out_restored)
```

```{r}
plot(out_urban)
```


```{r}
df_control <- as.data.frame(out_control)
#df_restored <- as.data.frame(out_restored)
#df_urban <- as.data.frame(out_urban)

fish_catch <- df_control %>% 
  dplyr::select(time, catch) %>% 
  dplyr::rename(Current = catch) %>% 
  mutate(Restored = df_restored$catch, Urban = df_urban$catch) %>% 
  gather('Current', 'Restored', 'Urban', key = "Scenario", value = "Mass")

#fish_dens <- df_control %>% 
  #dplyr::select(time, FB) %>% 
  #dplyr::rename(Current = FB) %>% 
  #mutate(Restored = df_restored$FB, Urban = df_urban$FB) %>% 
  #gather('Current', 'Restored', 'Urban', key = "Scenario", value = "Density")
  
```

```{r}
# Make a fish density (g/m^2) plot

n_plot <- ggplot(data = fish_catch, aes(x = time, y = Mass, col= Scenario)) +
  geom_line() +
  guides(fill=FALSE) +
  facet_wrap(~Scenario) +
  labs(title = "Modeled ‘Ama‘ama Catch From He'eia Fishpond", x = "Time (Years)", y = as.expression(bquote(Density (g/m^2)))) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5)

n_plot
```

```{r}
# Make a fish time series plot

catch_plot <- ggplot(data = fish_catch, aes(x = time, y = Mass, col= Scenario)) +
  geom_line() +
  guides(fill=FALSE) +
  facet_wrap(~Scenario) +
  labs(title = "Modeled ‘Ama‘ama Catch From He'eia Fishpond", x = "Time (Years)", y = as.expression(bquote( (g/m^2)))) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5)

catch_plot
```

```{r}
# Make a fish time series plot

catch_plot2 <- ggplot(data = fish_catch, aes(x = time, y = Mass, col= Scenario)) +
  geom_line() +
  guides(fill=FALSE) +
  labs(title = "Modeled ‘Ama‘ama Catch From He'eia Fishpond", x = "Time (Years)", y = as.expression(bquote(Density (g/m^3)))) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5)

catch_plot2
```

```{r}
# Make a fish time series plot

fish_dens_plot <- ggplot(data = fish_dens, aes(x = time, y = Density, col= Scenario)) +
  geom_line() +
  guides(fill=FALSE) +
  facet_wrap(~Scenario) +
  labs(title = "Modeled ‘Ama‘ama Density In He'eia Fishpond", x = "Time (Years)", y = as.expression(bquote(Density (g/m^3)))) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5)

fish_dens_plot
```

```{r}
# create a data frame of fishpond water column output to use in a figure
fp_df <- df_control %>% 
  dplyr::select(time, No_fp, Na_fp, Ni_fp, alg, fish, catch) %>% 
  gather('No_fp', 'Na_fp', 'Ni_fp', 'alg', 'fish', 'catch', key = "species", value = "concentration")

fp_df <- na.omit(fp_df)

neworder <- c('No_fp', 'Na_fp', 'Ni_fp', 'alg', 'fish', 'catch')

fp_df2 <- arrange(transform(fp_df,
             species=factor(species,levels=neworder)),species)
```


```{r}
# Six-panel figure of different state variables output from control conditions in the model

fp_plot <- ggplot(data = fp_df2, aes(x = time, y = concentration, col = species)) +
  facet_wrap(~species, scales = "free_y", 
                strip.position = "left", 
                labeller = as_labeller(c(alg = "Algae (g/m^2)", fish = "Fish (g/m^2)", Na_fp = "Ammonium-N (mg/L)", Ni_fp = "Nitrate-N (mg/L)", No_fp = "Organic N (mg/L)", catch = 'Caught Fish (kg)'))) +
  geom_line() +
  labs(title = "Model Outputs in He'eia Fishpond Water Column\n", x = "Time (Years)") +
  ylab(NULL) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), strip.background = element_blank(),
           strip.placement = "outside", legend.position = "none") 
fp_plot
```

```{r}
fp_water_df <- df_control %>% 
  dplyr::select(time, No_fp, Na_fp, Ni_fp) %>% 
  gather('No_fp', 'Na_fp', 'Ni_fp', key = "nitrogen", value = "concentration")

```


```{r}
# Make a nice ggplot with concentration of different nitrogen types in fishpond water
n_fp6_plot <- ggplot(data = fp_water_df, aes(x = time, y = concentration, col=nitrogen)) +
  geom_line() +
  guides(fill=FALSE) +
  labs(title = "Modeled Nitrogen Compounds in He'eia Fishpond Water Column\n Time Step = 6 h", x = "Time (Years)", y = "Concentration (mg/L)") +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(name="Nitrogen Types",
                         breaks=c("Na_fp", "Ni_fp", "No_fp"),
                        labels = c(as.expression(bquote(NH[4]^-1-N)), 
         as.expression(bquote(NO[3]^-1-N)), 
         "Organic N"))

n_fp6_plot
```


```{r}
# Three-panel figure of different nitrogen variables from control conditions in the fishpond water
fp_over20 <- fp_water_df %>% 
  filter(time >= 20 & time < 30)

fp_over20_plot <- ggplot(data = fp_over20, aes(x = time, y = concentration, col=nitrogen)) +
  facet_wrap(~nitrogen, scales = "free_y", 
                strip.position = "left", 
                labeller = as_labeller(c(No_fp = "Organic N (mg/L)", Na_fp = "Ammonium-N (mg/L)", Ni_fp = "Nitrate-N (mg/L)"))) +
  geom_line() +
  labs(title = "Model Outputs in He'eia Fishpond Water Column\n", x = "Time (Years)") +
  ylab(NULL) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), strip.background = element_blank(),
           strip.placement = "outside", legend.position = "none") 
fp_over20_plot
```


```{r}
fp_nit_late <- fp_water_df %>% 
  filter(nitrogen == "Ni_fp") %>% 
  filter(time >= 1 & time < 200)

nitrate_late = ggplot() + 
  geom_line(data = fp_nit_late, aes(x = time*365, y = concentration), col = 'green') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

nitrate_late
```

```{r}
fp_amm_late <- fp_water_df %>% filter(nitrogen == "Na_fp") %>% 
  filter(time >= 100 & time < 200)

amm_late = ggplot() + 
  geom_line(data = fp_amm_late, aes(x = time*365, y = concentration), col = 'red') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

amm_late
```

```{r, message=FALSE}
# Read in water quality data from 2014-2015 in He'eia Fishpond CSV
heeia_wq = read_csv('heeia_wq_data_2014_2015.csv') %>% 
  arrange(date) %>% 
  dplyr::select(date:turbidity)

```


```{r}
nitrate_data <- heeia_wq %>% 
  dplyr::select(date, statistic, nitrate) %>% 
  filter(statistic == "Mean" | statistic == "Stdv" | statistic == "N") %>% 
  spread(key = statistic, value = "nitrate") %>% 
  dplyr::rename(nit_obs = Mean, nit_sd = Stdv) %>% 
  mutate(nit_se = nit_sd / sqrt(N)) %>% 
  mutate(day_ct = 0, nit_var = nit_sd^2) %>% 
  na.omit()

  for (i in 2:length(nitrate_data$day_ct)){
    nitrate_data$day_ct[i] = nitrate_data$day_ct[i-1] + as.integer(nitrate_data$date[i]-nitrate_data$date[i-1])
  }

nitrate_data_shift <- nitrate_data %>% 
  mutate(day_shft = day_ct + 147) # shift the day counter from starting at 0 on the first day of data collection to instead start on the Julian Day of that year (147 for May 27). Then the day counter increases from there based on the day since sampling began
  #mutate(JD = yday(date))

ammonium_data <- heeia_wq %>% 
  dplyr::select(date, statistic, ammonia) %>% 
  filter(statistic == "Mean" | statistic == "Stdv" | statistic == "N") %>% 
  spread(key = statistic, value = "ammonia") %>% 
  dplyr::rename(amm_obs = Mean, amm_sd = Stdv) %>% 
  mutate(amm_se = amm_sd / sqrt(N)) %>% 
  mutate(day_ct = 0, amm_var = amm_sd^2) %>% 
  na.omit() %>% 
  filter(amm_obs != max(amm_obs)) # removed an outlier, could be related to a storm event. Could remove this line to add

  for (i in 2:length(ammonium_data$day_ct)){
    ammonium_data$day_ct[i] = ammonium_data$day_ct[i-1] + as.integer(ammonium_data$date[i]-ammonium_data$date[i-1])
  }

amm_data_shift <- ammonium_data %>% 
  mutate(day_shft = day_ct + 147) 
```



```{r}
fp_nitrate6 <- fp_water_df %>% 
  filter(nitrogen == "Ni_fp")

fp_ammonium6 <- fp_water_df %>% 
  filter(nitrogen == "Na_fp")
```

```{r}
# try to compare first year of modeled fishpond nitrate to the 2014-2015 fishpond water nitrate data
fp_nitrate_y1 <- fp_water_df %>% filter(nitrogen == "Ni_fp") %>% 
  filter(time >= 498 & time < 500)


nitate_combo_y1 = ggplot() + 
  geom_point(data = nitrate_data_shift, aes(x = day_shft, y = nit_obs),colour='blue', show.legend = T) +
  geom_errorbar(data = nitrate_data_shift, aes(x = day_shft, ymin=nit_obs-nit_sd, ymax=nit_obs+nit_sd), width=3, col='blue', show.legend = T) +
  geom_line(data = fp_nitrate_y1, aes(x = (time-498)*365, y = concentration), col = 'green') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))
  

nitate_combo_y1
```

```{r}
# compare to year 80 of modeled fishpond nitrate to the 2014-2015 fishpond water nitrate data
fp_nitrate_y20 <- fp_water_df %>% filter(nitrogen == "Ni_fp") %>% 
  filter(time >= 150 & time < 151.15)

  
nitate_combo_y20 = ggplot() + 
  geom_point(data = nitrate_data, aes(x = day_ct, y = nit_obs),colour='blue') +
  geom_errorbar(data = nitrate_data, aes(x = day_ct, ymin=nit_obs-nit_sd, ymax=nit_obs+nit_sd), width=3, col='blue') +
  geom_line(data = fp_nitrate_y20, aes(x = (time-150)*365, y = fp_nitrate_y20$concentration), col = 'green') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

nitate_combo_y20
```

```{r}
# compare to year 1 of modeled fishpond ammonium N to the 2014-2015 fishpond water ammonium data
fp_amm_y1 <- fp_water_df %>% filter(nitrogen == "Na_fp") %>% 
  filter(time >= 0 & time < 1.15)

ggplot(data = fp_amm_y1, aes(x = time, y = concentration)) +geom_line(col = 'red')

amm_combo_y1 = ggplot() + 
  geom_line(data = fp_amm_y1, aes(x = time*365, y = fp_amm_y1$concentration), col = 'red') + 
  geom_point(data = ammonium_data, aes(x = day_ct, y = amm_obs),col='magenta') +
  geom_errorbar(data = ammonium_data, aes(x = day_ct, ymin=amm_obs-amm_sd, ymax=amm_obs+amm_sd), width=3, col='magenta') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

amm_combo_y1
```


```{r}
# compare to year 80 of modeled fishpond ammonium N to the 2014-2015 fishpond water ammonium data
fp_amm_y20 <- fp_water_df %>% filter(nitrogen == "Na_fp") %>% 
  filter(time >= 498 & time < 500)


amm_combo_y20 = ggplot() +
  geom_line(data = fp_amm_y20, aes(x = (time-498)*365, y = fp_amm_y20$concentration), col = 'red') + 
  geom_point(data = amm_data_shift, aes(x = day_shft, y = amm_obs),col='magenta') +
  geom_errorbar(data = amm_data_shift, aes(x = day_shft, ymin=amm_obs-amm_sd, ymax=amm_obs+amm_sd), width=3, col='magenta') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

amm_combo_y20
```

Prepare the data to do the MSE/NSE analysis
```{r}
# need to average 4 times steps to get the average for that day. Assign a day number in integer form that can be compared against the days where there is data. Then make calculate MSE or some other metric for comparison

# create daily averages for nitrate concentrations modeled in the fishpond water
fp_nit_day <- data.frame(day_ct = seq(0,365*years),
                         nit_pred = 0)

for (i in 1:length(fp_nit_day$day_ct)){
  fp_nit_day$nit_pred[i] =  (fp_nitrate6$concentration[(4*i)-3]+fp_nitrate6$concentration[(4*i)-2]+ fp_nitrate6$concentration[(4*i)-1]+fp_nitrate6$concentration[4*i])/4
}

# create daily averages for ammonium concentrations modeled in the fishpond water
fp_amm_day <- data.frame(day_ct = seq(0,365*years),
                         amm_pred = 0)

for (i in 1:length(fp_amm_day$day_ct)){
  fp_amm_day$amm_pred[i] =  (fp_ammonium6$concentration[(4*i)-3]+fp_ammonium6$concentration[(4*i)-2]+ fp_ammonium6$concentration[(4*i)-1]+fp_ammonium6$concentration[4*i])/4
}

```

```{r}
# Merge the data frames 
amm_mse_df <- merge(ammonium_data, fp_amm_day, by = "day_ct")
nit_mse_df <- merge(nitrate_data, fp_nit_day, by = "day_ct")
```


```{r}
# checked this MSE function and it calculates the expected outcome
amm_mse = MSE(y_pred = amm_mse_df$amm_pred, y_true = amm_mse_df$amm_obs) 
nit_mse = MSE(y_pred = nit_mse_df$nit_pred, y_true = nit_mse_df$nit_obs)

# Nash-Sutcliffe Efficiency (NSE) is considered teh normalized MSE
# NSE = 1 - ( sum( (obs - sim)^2 ) / sum( (obs - mean(obs))^2 )
#Nash-Sutcliffe efficiencies range from -Inf to 1. Essentially, the closer to 1, the more accurate the model is. 
#-) NSE = 1, corresponds to a perfect match of modelled to the observed data. 
#-) NSE = 0, indicates that the model predictions are as accurate as the mean of the observed data, 
#-) -Inf < NSE < 0, indicates that the observed mean is better predictor than the model.

amm_var_mean = mean(amm_mse_df$amm_var)
nit_var_mean = mean(nit_mse_df$nit_var)

amm_nse = 1 - (amm_mse/amm_var_mean)
nit_nse = 1 - (nit_mse/nit_var_mean)

amm_nse # problematic that this is negative
nit_nse # probelmatic that this isn't that close to 1

# NSE seems to be used extensively in hydrologic models

# just by removing the outlier in the ammonium data set, the NSE went from  -1.138588 to 0.575
```

```{r}
amm_nse_plot <- ggplot(data = amm_mse_df) +
  geom_point(aes(x = day_ct, y = amm_obs),col='magenta') +
  geom_errorbar(aes(x = day_ct, ymin=amm_obs-amm_sd, ymax=amm_obs+amm_sd), width=3, col='magenta') +
  geom_point(aes(x = day_ct, y = amm_pred),col='blue') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

amm_nse_plot
```



Total food production from the He‘eia IAAS


Examine 10 year intervals for all land use scenarios. In the case of the restored scenario, the 10 years begin after 20 years of restoration are completed. Similarly for the urban land use scenario, the nitrogen loading used throughout the evaluated 10 year interval is constant, representing the loading from complete urbanization of the present.



Agriculture production

```{r}
current_catch <- df_control %>% 
  dplyr::select(time, catch)

current_catch_annual = as.numeric(rep(NA,years))

for(i in 1:years){
  
  catch_df_curr = current_catch %>% 
  dplyr::filter((i-1) <time & time <= (i))
  
  current_catch_annual[i] = mean(catch_df_curr$catch)
}

restored_catch <- df_restored %>% 
  dplyr::select(time, catch)

restored_catch_annual = as.numeric(rep(NA,years))

for(i in 1:years){
  
  catch_df_rest = restored_catch %>% 
  dplyr::filter((i-1) <time & time <= (i))
  
  restored_catch_annual[i] = mean(catch_df_rest$catch)
}

urban_catch <- df_urban %>% 
  dplyr::select(time, catch)

urban_catch_annual = as.numeric(rep(NA,years))

for(i in 1:years){
  
  catch_df_urb = urban_catch %>% 
  dplyr::filter((i-1) <time & time <= (i))
  
  urban_catch_annual[i] = mean(catch_df_urb$catch)
}

```

```{r}
tframe = seq(1,years) # specify the time frame of the years of interest

# Use the mean of the range of the yield values for the three different crops
banana_yield = 1.49 # mean of 0.58-2.39 kg/yr from Bremer, et al (2018)
breadfruit_yield = 1.32 # mean of 0.20-2.43 kg/yr from Bremer, et al (2018)
taro_yield = 1.46 # mean of 1.12-1.79 kg/yr from Bremer, et al (2018)

banana_annual_add = 2428 # [m^2/yr]
breadfruit_annual_add = 4856 # [m^2/yr]
taro_annual_add = 6475 # [m^2/yr]

banana_area_current = 0 # [m^2] current banana production area
breadfruit_area_current = 0 # [m^2] current breadfruit production area
taro_area_current = 7300 # [m^2] current taro production area

# current land use mass [kg]
banana_current_mass = rep(banana_area_current*banana_yield, years)
breadfruit_current_mass = rep(breadfruit_area_current*breadfruit_yield, years)
taro_current_mass = rep(taro_area_current*taro_yield, years)

# restored land use mass [kg]
banana_rest_mass = (banana_area_current + banana_annual_add*tframe)*banana_yield
breadfruit_rest_mass = (breadfruit_area_current + breadfruit_annual_add*tframe)*breadfruit_yield
taro_rest_mass = (taro_area_current + taro_annual_add*tframe)*taro_yield

# urban land use mass [kg]
banana_urban_mass = rep(0, years)
breadfruit_urban_mass = rep(0, years)
taro_urban_mass = rep(0, years)

Crop=c(rep("Taro", 3*(years)), rep("Banana", 3*(years)), rep("Breadfruit", 3*(years)), rep("Fish", 3*(years)))
Scenario =rep(c("Current", "Restored", "Urban") , each = years, times = 4)
Mass=c(taro_current_mass/1000, taro_rest_mass/1000, taro_urban_mass/1000, banana_current_mass/1000, banana_rest_mass/1000, banana_urban_mass/1000, breadfruit_current_mass/1000, breadfruit_rest_mass/1000, breadfruit_urban_mass/1000, current_catch_annual/1000, restored_catch_annual/1000, urban_catch_annual/1000) # convert all the kg masses to metric tons
Time = as.numeric(rep(tframe, times = 12))

crop_data=data.frame(Time,Crop,Scenario,Mass)

```


 
```{r}
# Stacked
crop_stacked <- ggplot(crop_data, aes(fill = Crop, y=Mass, x=Time)) + 
    geom_bar(stat="identity") +
    facet_wrap(~Scenario) +
    #scale_fill_discrete_qualitative(palette = "Cold") +
    guides(fill=guide_legend()) +
    labs(y = "Metric Tons", x = "Years") +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) +
    theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5, strip.background = element_blank())

crop_stacked # difficult to see the mass from fish catch
```

```{r}
crop_curr <- crop_data %>% 
  filter(Scenario == "Current") %>% 
  dplyr::select(Time, Mass) %>%
  colSums()

crop_rest <- crop_data %>% 
  filter(Scenario == "Restored") %>% 
  dplyr::select(Time, Mass) %>%
  colSums()

crop_urban <- crop_data %>% 
  filter(Scenario == "Urban") %>% 
  dplyr::select(Time, Mass) %>%
  colSums()


```



What if you were instead going to do a time series of 10 or 20 years from current conditions to either restored or urban conditions and you add a linear proportion of what the loading is supposed to increase by with each successive year?


```{r}
# extract the last value for the fish density in each land use scenario and multiply by He‘eia Fishpond's volume to get the mass in grams and convert to metric tons
current_fish <- df_control %>% 
  dplyr::select(time, catch) %>% 
  filter(time > 9) %>% 
  colSums() %>% 
  as.data.frame()

fish_ctrl = current_fish$.[2]*A_fp/10^6
 # check if it should be V_fp or A_fp at the end
restored_fish <- df_restored %>% 
  dplyr::select(time, catch) %>% 
  filter(time > 9) %>% 
  colSums() %>% 
  as.data.frame()

fish_rst = restored_fish$.[2]*A_fp/10^6

urban_fish <- df_urban %>% 
  dplyr::select(time, catch) %>% 
  filter(time > 9) %>% 
  colSums() %>% 
  as.data.frame()

fish_urb = urban_fish$.[2]*A_fp/10^6
```