---
title: "He’eia Model - Stream PFR"
author: "Nakoa Farrant"
date: "5/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(deSolve)
library(ggplot2)
library(schoolmath)
library(tidyverse)
library(TropFishR)
require(MASS)
require(dplyr)
library(colorspace)
library(MLmetrics)
```

```{r}
# The below values are characteristic of Mugil cephalus (striped mullet, ‘ama‘ama) as defined in the S.I. of Cheung, et al. 2013, who derived these values from data that is freely available at fishbase.org
Linf_mugil = 100 # asymptotic length for Mugil cephalus, [cm]
K_mugil = 0.27 # von Bertalanffy growth parameter, [1/year]
a_mugil = 0.032 # a for overall M.cephauls on fishbase is 0.0085 but this a is specific to M. cephalus in Hawai'i estuaries. length-weight conversion parameter from FishBase
b_mugil = 3.0 # length-weight conversion parameter from FishBase
Winf_mugil = a_mugil*Linf_mugil^b_mugil # asymptotic weight using length-weight conversion [grams]

# For Mugil cephalus (striped mullet), natural mortality can be empirically derived using the method developed by Then, et al. 2015
M_mugil = M_empirical(Linf = Linf_mugil, K_l = K_mugil, method = "Then_growth") # [1/year]
```

# 6-Hour Time Step: Model Equations
Realized that running into issues with the stream and tide specifications that are determined based on the value of t and need to be adjusted based on the time step being taken
```{r}

fish_func6 <- function(t, state6, pars6) {
  with(as.list(c(state6, pars6)), {
    
    # create an integer counter variable, time is otherwise in 6 hr intervals
    step = t*365*4 
    
    # temperature in ºC
    T_max = 35.2 # maximum in the summer
    T_min = 20.5 # minimum in the winter
    T1 = (T_max - T_min)/2 * (sin(2*pi*step - pi/2) + 1) + T_min
    
    #k_fd = (0.0256 + 0.0123*T1)/4 # temperature-driven fish mortality rate [1/6hr]
    
    # define flow terms- m^3/d
    Q_st = 3500/4 # average He‘eia stream flow conditions
    
    # There are 4 tidal cycles every day that alternate between flood and ebb tides at each 6 hour time     step. Additionally, roughly every 8 days there is a switch from spring tide to neap tide based on       the moon cycle (spring when full or new moon, neap during quarter moon phases). This results in         desired shifts from spring to neap every ~32 tide shifts (8 days * 4 tides/day)
    
    # Every lunar cycle (approximately a month), there are two sets of spring and two sets of neap tides     # which totals to 128 tidal cycles. By using the modulo operator based on 128 tidal cycles, you can     # get all of the spring and neap cycles and all of the flood and ebb cycles in each month which will     # repeat each month.
    if (step %% 128 < 32){
      # arbitrarily start with the first 32 spring tide cycles
      if (as.integer(step) %% 2 == 0){
        # alternate between spring flood and spring ebb based on being an even or odd step
        Q_tide = sp_fl
      }
      else{
        Q_tide = sp_ebb
      }
    }
    else if (step %% 128 > 31 & step %% 128 < 64){
      # shift to  32 neap tides
      if(as.integer(step) %% 2 == 0){
        # alternate between neap flood and neap ebb based on being an even or odd step
        Q_tide = np_fl
      }
      else{
        Q_tide = np_ebb
      }
    }
    else if (step %% 128 > 63 & step %% 128 < 96){
      # shift back to spring tides
      if(as.integer(step) %% 2 == 0){
        # alternate between spring flood and spring ebb based on being an even or odd step
        Q_tide = sp_fl
      }
      else{
        Q_tide = sp_ebb
      }
    }
    else{
      # shift back to neap tides
      if(as.integer(step) %% 2 == 0){
        # alternate between neap flood and neap ebb based on being an even or odd step
        Q_tide = np_fl
      }
      else{
        Q_tide = np_ebb
      }
    }
    
     
    # Specify temporal variability in He'eia stream flow rate based on mean
    # measurements from July 2014-July 2015
    # m^3 / 6 hours
    if (step %% 365 < 10) {
        # End of July 2014
        Q_st = 8860/4
    }
    else if (step %% 365 > 9 && step %% 365 < 40) 
    {
        # Aug 2014
        Q_st = 4620/4
    }
    else if (step %% 365 > 39 && step %% 365 < 68) 
    {
        # Sept 2014
        Q_st = 4650/4
    }
    else if (step %% 365 > 67 && step %% 365 < 99) 
    {
        # Oct 2014
        Q_st = 6560/4 
    }
    else if (step %% 365 > 98 && step %% 365 > 129) 
    { 
        # Nov 2014
        Q_st = 4970/4 
    }
    else if (step %% 365 > 128 && step %% 365 < 160) 
    {
       # Dec 2014 
       Q_st = 4580/4
    }
    else if (step %% 365 > 159 && step %% 365 < 191) 
    {
       # Jan 2015 
       Q_st = 4130/4 
    }
    else if (step %% 365 > 190 && step %% 365 < 219) 
    { 
       # Feb 2015 
       Q_st = 3990/4 
   }
    else if (step %% 365 > 218 && step %% 365 < 250) 
    {
       # March 2015 
       Q_st = 3940/4 
    }
    else if (step %% 365 > 249 && step %% 365 < 280) 
    {
       # April 2015
       Q_st =  3840/4 
    }
    else if (step %% 365 > 279 && step %% 365 < 312) 
    {
       # May 2015
       Q_st =  3600/4
    }
    else if (step %% 365 > 311 && step %% 365 < 342) 
    {
       # June 2015
       Q_st =  4160/4
    } 
    else if (step %% 365 > 341 && step %% 365 < 366) 
    {
      # remainder of July 2015
       Q_st = 3770/4 
    }
    else {
       Q_st = 3500/4 # average stream flow during this period
    }

    
    W_No = LU_No*Q_st # loading of organic N runoff based on the evaluated land use organic N concentration in the He‘eia Wetland [g/6hr] = g/m^3 * m^3/6hr
    W_Na = LU_Na*Q_st # loading of organic N runoff based on the evaluated land use ammonium N concentration in the He‘eia Wetland [g/6hr]
    W_Ni = LU_Ni*Q_st # loading of organic N runoff based on the evaluated land use nitrate N concentration in the He‘eia Wetland [g/6hr]
    
  # Stream organic N
  dNo_st <- (W_No- Q_st*No_st+(-k_oa*No_st + k_ao*Na_st)*V_st)/V_st
  #dNo_st <- LU_No*exp(-k_oa/Q_st*l_st) + LU_Na*exp(k_ao/Q_st*l_st)
  # Stream ammonium
  dNa_st <- (W_Na - Q_st*Na_st+(k_oa*No_st-(k_ao+k_nit+k_av)*Na_st + k_nit_red*Ni_st)*V_st)/V_st
  #dNa_st <- LU_Na*exp((-(k_ao+k_nit+k_av))/Q_st*l_st) + LU_No*exp(k_oa/Q_st*l_st) + LU_Ni*exp(k_nit_red/Q_st*l_st)
  # Stream nitrate    
  dNi_st <- (W_Ni- Q_st*Ni_st + (k_nit*Na_st-k_nit_red*Ni_st)*V_st)/V_st
  #dNi_st <- LU_Ni*exp(-k_nit_red/Q_st*l_st) + LU_Na*exp(k_nit/Q_st*l_st)
  # Fishpond water organic N
  dNo_fp <- ((Q_st*No_st+((-k_oa-k_s)*No_fp+k_ao*Na_fp+k_r*No_sd)*V_fp + Q_tide*(No_ocean-No_fp)))/V_fp
  # Fishpond water ammonium    
  dNa_fp <- (Q_st*Na_st + (k_oa*No_fp - (k_ao+k_nit+k_av+k_s)*Na_fp + k_r*Na_sd)*V_fp+Q_tide*
             (Na_ocean-Na_fp)-((r_NChl*k_g20*((Na_fp+Ni_fp)/(k_SN+No_fp+Ni_fp))+r_NChl*k_ra)*alg))/V_fp
  # Fishpond water nitrate    
  dNi_fp <- (Q_st*Ni_st + (k_nit*Na_fp-k_s*Ni_fp+k_r*Ni_sd)*V_fp + Q_tide*(Ni_ocean-Ni_fp)-
             ((r_NChl*k_g20*((Na_fp+Ni_fp)/(k_SN+Na_fp+Ni_fp)) + r_NChl*k_ra)*alg))/V_fp
  # Sediment organic N    
  dNo_sd <- ((k_s*No_fp-(k_oa+k_b+k_r)*No_sd + k_ao*Na_sd)*V_sed)/V_sed
  # Sediment ammonium
  dNa_sd <- ((k_s*Na_fp + k_oa*No_sd-(k_ao+k_nit_sed+k_r+k_b)*Na_sd + k_nit_red*Ni_sd)*V_sed)/V_sed
  # Sediment nitrate    
  dNi_sd <- ((k_s*Ni_fp+k_nit_sed*Na_sd-(k_denit+k_nit_red+k_r+ k_b)*Ni_sd)*V_sed)/V_sed
  # Algae density
  dalg <- ((r_NChl*k_g20*((Na_fp+Ni_fp)/(k_SN+Na_fp+Ni_fp)) -k_ra -((Q_st-Q_tide)/V_fp)))*
           alg - k_ac*fish
  # Fish density
  if(step < 365*4)
    catch = 0
  else
    catch = k_catch
  
  dfish <- (k_fg - k_fd - catch)*fish # question the parentheses location from a units perspective
  #dfish <- k_fg*fish*(1-fish/carrying)
dcatch = catch*fish
  return(list(c(No_st = dNo_st, Na_st = dNa_st, Ni_st = dNi_st, No_fp = dNo_fp, Na_fp = dNa_fp, Ni_fp = dNi_fp, No_sd = dNo_sd, Na_sd = dNa_sd, Ni_sd = dNi_sd, alg = dalg, fish = dfish, catch = dcatch)))

})
}
```

# Model Specification
```{r}
years = 50
t <- seq(0, years, by = 1/365/4) # 100 years with a 6 hour time step

# define physical parameters of ecosystem 
h_st = 0.5 # [m] average depth of stream 
l_st = 700 # [m] length of interest for He‘eia Stream
w_st = 15 # [m] approximate width of stream

A_fp = 360000 # [m^2] area of fishpond
h_fp = 0.4 # [m] average depth of fishpond water column [m]

A_sed = 360000 # [m^2] area of sediment below fishpond water column
h_sed = 0.005 # [m] depth of sediment of interest in meters

# define volumes of compartments - m^3
V_st = h_st*l_st*w_st # stream
V_fp = A_fp*h_fp # fishpond water
V_sed = A_sed*h_sed # fishpond sediment

pars_control <- c(
      LU_No = 1.8, # [g/m^3] organic N concentration in wetland for current land use
      LU_Na = 0.07, # [g/m^3] ammonium N concentration in wetland for current land use
      LU_Ni = 0.07, # [g/m^3] nitrate N concentration in wetland for current land use
      
      carrying = 56, # [g/m^2] carrying capacity for fish in He‘eia Fishpond, 500 pounds per acre
      
      k_oa = 0.01/4, # [1/6hr] conversion of organic N to ammonium N
      k_ao = 0.1/4, # [1/6hr] conversion of ammonium N to organic N 
      k_nit = 0.5/4, # [1/6hr] nitrification in water column
      k_nit_red = 0.1/4, # [1/6hr] reduction of nitrate N to ammonium N
      k_av = 0.7/4, # [1/6hr] ammonium to ammonia volatilization rate
      k_nit_sed = 0.05/4, # [1/6hr] nitrification rate in the sediment
      k_denit = 0.01/4, # [1/6hr] denitrification rate constant
      
      # Fishpond water-sediment exchange rate constants
      k_r = 0.4/4, # [1/6hr] resuspension rate
      k_s = 0.8/4, # [1/6hr] settling rate, used to be 0.05 1/day. If going to make it this high, maybe it should just be related to phytoplankton settling rate due to their size. This is from Burdoff and Lorenzen
      k_b = 0.001/4, # [1/6hr] burial removal rate [1/6hr] 
      
      # fish and algae rates and constants
      r_NChl = 0.2, # [g N/g Chla] ratio of nitrogen to Chla in algae
      k_ra = 0.025/4, # [1/6hr] algae respiration rate
      k_SN = 0.015, # [gN/m^3] half-saturation constant for N uptake. Chapra says 10-15 mgN/m^3. Previously had this at 150 mg/m^3
      k_g20 = 1.5/4, # [m] algae growth rate constant at 20ºC
      #k_fg = 0.015/4, # rate of fish growth due to algae consumption [1/6hr]
      # K for mullet is 0.27 1/yr -> 0.0007 1/day, this is lower than the current growth rate but the current growth rate also seems particularly low
      k_ac = 0.08, # expect M. cephalus to consume ~8% of it's body weight in food, De Silva & Perera 
      k_fg = 0.27/365/4, # [1/6hr]
      #k_fg = K_mugil/365/4, # rate of fish growth [1/6hr], K_mugil = 0.27
      # based on doubling time of 1.4-4.4 years, r_growth = ln(2)/1.4 = 0.495 /yr to ln(2)/4.4 = 0.158 /yr, prior r (whatever that means) is 0.51
      k_fd = M_mugil/365/4, # [1/6hr] rate of fish death from empirical Then formula, M_mugil = 0.346
      k_catch = M_mugil/365/4, # could be 0.343 based on Mexico study. file:///Users/Nakoa/Downloads/2016GallardoetalfisherydeMugelcephalusJMBO-5-151.pdf
      # fishbase also says you generally want Exploitation E to be about 0.5 which is F/Z = F/(F+M) so M and F should be approximately equal ideally for MSY
      # From Barman, Jana, Garg, Bhatnagar, https://link.springer.com/article/10.1007/s10499-004-2479-5
      # growth rate for M. cephalus in salinity 20 ppt 0.69 g/d or 0.67 g/d in 25 ppt
      # SGR --specific growth rate of weight-- (% BW (in grams) per day) about 4% per day
      k_birth = 0.35/365/4, # [1/6hr]
      
      # nitrogen concentrations in ocean - units g/m^3
      No_ocean = 0.001,
      Na_ocean = 0.04, 
      Ni_ocean = 0.02,
      
      # water volume flux rate (m^3/6hr)
      sp_fl= 191660/4,
      sp_ebb = -174880/4,
      np_fl = 141384/4,
      np_ebb = -159938/4
)


pars_restored <- c(
      LU_No = 6.2, # [g/m^3] organic N concentration in wetland for current land use
      LU_Na = 0.25, # [g/m^3] ammonium N concentration in wetland for current land use
      LU_Ni = 0.25, # [g/m^3] nitrate N concentration in wetland for current land use
      
      carrying = 56, # carrying capacity for fish in He‘eia Fishpond, 500 pounds per acre
      
      k_oa = 0.01/4, # conversion of organic N to ammonium N [1/6hr]
      k_ao = 0.1/4, # conversion of ammonium N to organic N [1/6hr]
      k_nit = 0.5/4, # nitrification in water column [1/6hr]
      k_nit_red = 0.1/4, # reduction of nitrate N to ammonium N [1/6hr]
      k_av = 0.7/4, # ammonium to ammonia volatilization rate [1/6hr]
      k_nit_sed = 0.05/4, # nitrification rate in the sediment [1/6hr]
      k_denit = 0.01/4, # denitrification rate constant [1/6hr]
      
      # Fishpond water-sediment exchange rate constants
      k_r = 0.1/4, # resuspension rate [1/6hr]
      k_s = 0.8/4, # settling rate [1/6hr], used to be 0.05 1/day. If going to make it this high, maybe it should just be related to phytoplankton settling rate due to their size. This is from Burdoff and Lorenzen
      k_b = 0.001/4, # burial removal rate [1/6hr] 
      
      # fish and algae rates and constants
      r_NChl = 0.2, # ratio of nitrogen to Chla in algae [g N/g Chla]
      k_ra = 0.025/4, # algae respiration rate [1/6hr]
      k_SN = 0.015, # half-saturation constant for N uptake [gN/m^3], Chapra says 10-15 mgN/m^3. Previously had this at 150 mg/m^3
      k_g20 = 1.5/4, # algae growth rate constant at 20 degrees C [1/6hr]
      #k_fg = 0.015/4, # rate of fish growth due to algae consumption [1/6hr]
      # K for mullet is 0.27 1/yr -> 0.0007 1/day, this is lower than the current growth rate but the current growth rate also seems particularly low
      k_ac = 0.0012/4, # rate of algae consumption by fish [1/6hr]
      k_fg = K_mugil/365/4, # rate of fish growth [1/6hr], K_mugil = 0.27
      k_fd = M_mugil/365/4, # rate of fish death from empirical Then formula [1/6hr], M_mugil = 0.346
      k_catch = 0.05/365/4,
      k_birth = 0.35/365/4,
      
      # nitrogen concentrations in ocean - units g/m^3
      No_ocean = 0.001,
      Na_ocean = 0.04, 
      Ni_ocean = 0.02,
      
      # water volume flux rate (m^3/6hr)
      sp_fl= 191660/4,
      sp_ebb = -174880/4,
      np_fl = 141384/4,
      np_ebb = -159938/4
)

pars_urban <- c(
      LU_No = 13, # [g/m^3] organic N concentration in wetland for current land use
      LU_Na = 0.53, # [g/m^3] ammonium N concentration in wetland for current land use
      LU_Ni = 0.53, # [g/m^3] nitrate N concentration in wetland for current land use
      
      carrying = 56, # [g/m^2], carrying capacity for fish in He‘eia Fishpond, 500 pounds per acre from Keala loko i‘a doc
      
      k_oa = 0.01/4, # conversion of organic N to ammonium N [1/6hr]
      k_ao = 0.1/4, # conversion of ammonium N to organic N [1/6hr]
      k_nit = 0.5/4, # nitrification in water column [1/6hr]
      k_nit_red = 0.1/4, # reduction of nitrate N to ammonium N [1/6hr]
      k_av = 0.7/4, # ammonium to ammonia volatilization rate [1/6hr]
      k_nit_sed = 0.05/4, # nitrification rate in the sediment [1/6hr]
      k_denit = 0.01/4, # denitrification rate constant [1/6hr]
      
      # Fishpond water-sediment exchange rate constants
      k_r = 0.1/4, # resuspension rate [1/6hr]
      k_s = 0.8/4, # settling rate [1/6hr], used to be 0.05 1/day. If going to make it this high, maybe it should just be related to phytoplankton settling rate due to their size. This is from Burdoff and Lorenzen
      k_b = 0.001/4, # burial removal rate [1/6hr] 
      
      # fish and algae rates and constants
      r_NChl = 0.2, # ratio of nitrogen to Chla in algae [g N/g Chla]
      k_ra = 0.025/4, # algae respiration rate [1/6hr]
      k_SN = 0.015, # half-saturation constant for N uptake [gN/m^3], Chapra says 10-15 mgN/m^3. Previously had this at 150 mg/m^3
      k_g20 = 1.5/4, # algae growth rate constant at 20 degrees C [1/6hr]
      #k_fg = 0.015/4, # rate of fish growth due to algae consumption [1/6hr]
      # K for mullet is 0.27 1/yr -> 0.0007 1/day, this is lower than the current growth rate but the current growth rate also seems particularly low
      k_ac = 0.0012/4, # rate of algae consumption by fish [1/6hr]
      k_fg = 1.3/365/4,
      #k_fg = K_mugil/365/4, # rate of fish growth [1/6hr], K_mugil = 0.27
      k_fd = M_mugil/365/4, # rate of fish death from empirical Then formula [1/6hr], M_mugil = 0.346
      k_catch = 0.05/365/4,
      k_birth = 0.35/365/4,
      
      # nitrogen concentrations in ocean - units g/m^3
      No_ocean = 0.001,
      Na_ocean = 0.04, 
      Ni_ocean = 0.02,
      
      # water volume flux rate (m^3/6hr)
      sp_fl= 191660/4,
      sp_ebb = -174880/4,
      np_fl = 141384/4,
      np_ebb = -159938/4
)
# there will be seasonal variability in the baseline nitrogen forms present in the stream, fishpond and sediment based on tides, timing of nitrogen applications to the lo‘i, precipitation

# specify initial concentrations of nitrogen forms (g/m^3) and for algae and fish (g/m^2)

# Information on nitrogen output from Kako‘o taro agriculture from Bremer, et al 2018: total N is 0.16 
# (+/- 0.08) g/m^3 with nitrate N 0.006 (+/- 0.014) g/m^3 and ammonia N at 0.006 (+/- 0.006) g/m^3
# these represent an order of magnitude smaller ammonia and nitrate N composition from before

# Nitrogen in FP water column (no measures of organic N):
# summer mean:
# nitrate = 0.66 g/m^3
# ammonium = 0.02 g/m^3

# winter mean:
# nitrate = 0.69 g/m^3
# ammonium = 0.05 g/m^3

# low measured levels of nitrite (0.01 and 0.02 g/^3 in winter and summer, respectively)

# Sediment nitrogen concentrations based on the Briggs, et al paper from 2013 on coastal endmembers

# conversion factor for NH4-N: 1 mg NH4/l = 55.4 umol NH4/l

# In the first 1cm of sediment:
# ammonium (in terrogenous near egret island): 0.54 g/m^3 (~30 uM)
# ammonium (in carbonate sediment in more central part of pond): 0.63 g/m^3 (~35 uM)

# 0.59 g/m^3 average

# organic N: 

# waiting on organic N in FP and sediment, nitrate in sediment.


state6 <- c(
           No_st = 0.148, # [g/m^3] stream initial organic N
           Na_st = 0.006, # [g/m^3] stream initial ammonium
           Ni_st = 0.006, # [g/m^3] stream initial nitrate
           No_fp = 0.1, # [g/m^3] fishpond initial organic N, estimated from Charles Young thesis
           Na_fp = 0.035, # [g/m^3] fishpond initial ammonium
           Ni_fp = 0.675, # [g/m^3] fishpond initial nitrate
           No_sd = 0.1, # [g/m^3] fishpond sediment initial organic N, estimated from Young thesis
           Na_sd = 0.6, # [g/m^3] fishpond sediment initial ammonium, sediment is a source for ammonium according to Hargreaves (1998)
           Ni_sd = 0.01, # [g/m^3]  fishpond sediment initial nitrate, sediment is a sink for nitrate according to Hargreaves (1998).
           alg = 0.03, # [g algae/m^2] initial algae concentration based on average of data from Frank, et al (0.0015 g Chl a/m^3) to g/m^2 algae. 0.0015 g Chla/m^3 / 0.02 g algae/g Chla* 0.4 m
           # range of conversion factors [g alg/g Chla] from Chapra 0.005-0.025. Would be 0.12 g/m^2 algae if used 0.005
           # Using conversion factor from Chapra
           fish = 10, # [g/m^2] initial ‘ama‘ama  density
           catch = 0) 

# According to Mohlenkamp et al. 2019, from 2006-2009, Paepae o He'eia produced ~1.2 metric tons of moi (Pacific Threadfin) which is about 400 kg per year

# historical estimates for this fishpond max out around 15 metric tons (15,000 kg) with an asymptotic weight of 8.5 kg per striped mullet this would amount to ~1750 fully grown striped mullet.
# According to FishBase, the common length is 50 cm which equates to a weight closer to 1 kg based on the length-weight conversion factor.  This would require closer to 15,000 fish to reach capacity

# hypothetically stock with 3 cm fingerlings that have a mass of ~230 grams each based on length-weight conversion. These fish would be an estimated 1.5 years old based on the VBGF curve and the estimated weight. If we stock the fishpond with 5,000 fingerlings, the inital mass in the fishpond is 1.15*10^6 grams (1.1 metric tons). Initial fish density [g/m^3] 1150000 g/360000m^3 = 3.19 g/m^3 = 3.2 g/m^3

# volume = 144000m^3. 1150000 g/144000m^3 = 7.986 g/m^3 = 8 g/m^3

# fish base indicates sexual maturity at 3-4 years for Mugil cephalus. At the end of 4 years, fish are estimated to be ~3.2 kg each, which would translate to 15 metric tons

# Biswas, et al. 2012 (https://www.sciencedirect.com/science/article/pii/S0044848612000658) suggests a stocking density of M. cephalus fry (~0.17 g in size) at 15,000 fry/hectare. He'eia fishpond is 36 ha. This is 91,800 grams of M. cephalus stocked in the fishpond, which is close to what was suggested above with slightly larger fish

# 1150000 g/144000m^3 = 7.986 g/m^3
# fish carrying capacity at 500 pounds per acre -> 56 g/m^2



```


```{r}
out_control <- ode(y = state6, times = t, func = fish_func6, parms = pars_control, method = "euler")
out_restored <- ode(y = state6, times = t, func = fish_func6, parms = pars_restored, method = "euler")
out_urban <- ode(y = state6, times = t, func = fish_func6, parms = pars_urban, method = "euler")
# the method used to solve the ODE matters for the results
```

```{r}
plot(out_control)
```

```{r}
plot(out_restored)
```

```{r}
plot(out_urban)
```


```{r}
df_control <- as.data.frame(out_control)
df_restored <- as.data.frame(out_restored)
df_urban <- as.data.frame(out_urban)

fish_catch <- df_control %>% 
  dplyr::select(time, catch) %>% 
  dplyr::rename(Current = catch) %>% 
  mutate(Restored = df_restored$catch, Urban = df_urban$catch) %>% 
  gather('Current', 'Restored', 'Urban', key = "Scenario", value = "Density")
  
```


```{r}
# Make a fish time series plot

fish_plot <- ggplot(data = fish_catch, aes(x = time, y = Density, col= Scenario)) +
  geom_line() +
  guides(fill=FALSE) +
  labs(title = "Modeled Mugil cephalus Density in He'eia Fishpond", x = "Time (Years)", y = as.expression(bquote(Density (g/m^3)))) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5)

fish_plot
```

```{r}
# create a data frame of fishpond water column output to use in a figure
fp_df <- df_control %>% 
  dplyr::select(time, No_fp, Na_fp, Ni_fp, alg, fish) %>% 
  gather('No_fp', 'Na_fp', 'Ni_fp', 'alg', 'fish', key = "species", value = "concentration")

neworder <- c('No_fp', 'Na_fp', 'Ni_fp', 'alg', 'fish')

fp_df2 <- arrange(transform(fp_df,
             species=factor(species,levels=neworder)),species)
```


```{r}

fp_plot <- ggplot(data = fp_df2, aes(x = time, y = concentration, col = species)) +
  facet_wrap(~species, scales = "free_y", 
                strip.position = "left", 
                labeller = as_labeller(c(alg = "Algae (mg Chla/L)", fish = "Fish (g/m^2)", Na_fp = "Ammonium-N (mg/L)", Ni_fp = "Nitrate-N (mg/L)", No_fp = "Organic N (mg/L)"))) +
  geom_line() +
  labs(title = "Model Outputs in He'eia Fishpond Water Column\n", x = "Time (Years)") +
  ylab(NULL) +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), strip.background = element_blank(),
           strip.placement = "outside", legend.position = "none") 
fp_plot
```


```{r}
# Make a nice ggplot with desired content
fp_water_df <- df_control %>% 
  dplyr::select(time, No_fp, Na_fp, Ni_fp) %>% 
  gather('No_fp', 'Na_fp', 'Ni_fp', key = "nitrogen", value = "concentration")

n_fp6_plot <- ggplot(data = fp_water_df, aes(x = time, y = concentration, col=nitrogen)) +
  geom_line() +
  guides(fill=FALSE) +
  labs(title = "Modeled Nitrogen Compounds in He'eia Fishpond Water Column\n Time Step = 6 h", x = "Time (Years)", y = "Concentration (mg/L)") +
  theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(name="Nitrogen Types",
                         breaks=c("Na_fp", "Ni_fp", "No_fp"),
                        labels = c(as.expression(bquote(NH[4]^-1-N)), 
         as.expression(bquote(NO[3]^-1-N)), 
         "Organic N"))

n_fp6_plot
```


```{r, message=FALSE}
# Read in water quality data from 2014-2015 in He'eia Fishpond CSV
heeia_data = read_csv('heeia_wq_data_2014_2015.csv') %>% 
  arrange(date) %>% 
  dplyr::select(date:turbidity)

```


```{r}
nitrate_data <- heeia_data %>% 
  dplyr::select(date, statistic, nitrate) %>% 
  filter(statistic == "Mean" | statistic == "Stdv" | statistic == "N") %>% 
  spread(key = statistic, value = "nitrate") %>% 
  dplyr::rename(nit_obs = Mean, nit_sd = Stdv) %>% 
  mutate(nit_se = nit_sd / sqrt(N)) %>% 
  mutate(day_ct = 0, nit_var = nit_sd^2) %>% 
  na.omit()

  for (i in 3:length(nitrate_data$day_ct)){
    nitrate_data$day_ct[i-1] = nitrate_data$day_ct[i-2] + as.integer(nitrate_data$date[i]-nitrate_data$date[i-1])
  }

ammonium_data <- heeia_data %>% 
  dplyr::select(date, statistic, ammonia) %>% 
  filter(statistic == "Mean" | statistic == "Stdv" | statistic == "N") %>% 
  spread(key = statistic, value = "ammonia") %>% 
  dplyr::rename(amm_obs = Mean, amm_sd = Stdv) %>% 
  mutate(amm_se = amm_sd / sqrt(N)) %>% 
  mutate(day_ct = 0, amm_var = amm_sd^2) %>% 
  na.omit() %>% 
  filter(amm_obs != max(amm_obs)) # removed an outlier, could be related to a storm event. Could remove this line to add

  for (i in 3:length(ammonium_data$day_ct)){
    ammonium_data$day_ct[i-1] = ammonium_data$day_ct[i-2] + as.integer(ammonium_data$date[i]-ammonium_data$date[i-1])
  }
```



```{r}
fp_nitrate6 <- fp_water_df %>% 
  filter(nitrogen == "Ni_fp")

fp_ammonium6 <- fp_water_df %>% 
  filter(nitrogen == "Na_fp")
```

```{r}
# try to compare first year of modeled fishpond nitrate to the 2014-2015 fishpond water nitrate data
fp_nitrate_y1 <- fp_water_df %>% filter(nitrogen == "Ni_fp") %>% 
  filter(time >= 0 & time < 1.15)


nitate_combo_y1 = ggplot() + 
  geom_point(data = nitrate_data, aes(x = day_ct, y = nit_obs),colour='blue') +
  geom_errorbar(data = nitrate_data, aes(x = day_ct, ymin=nit_obs-nit_sd, ymax=nit_obs+nit_sd), width=3, col='blue') +
  geom_line(data = fp_nitrate_y1, aes(x = time*365, y = concentration), col = 'green') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))
  

nitate_combo_y1
```

```{r}
# compare to year 80 of modeled fishpond nitrate to the 2014-2015 fishpond water nitrate data
fp_nitrate_y20 <- fp_water_df %>% filter(nitrogen == "Ni_fp") %>% 
  filter(time >= 19 & time < 20.15)

  
nitate_combo_y20 = ggplot() + 
  geom_point(data = nitrate_data, aes(x = day_ct, y = nit_obs),colour='blue') +
  geom_errorbar(data = nitrate_data, aes(x = day_ct, ymin=nit_obs-nit_sd, ymax=nit_obs+nit_sd), width=3, col='blue') +
  geom_line(data = fp_nitrate_y20, aes(x = (time-19)*365, y = fp_nitrate_y20$concentration), col = 'green') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

nitate_combo_y20
```

```{r}
# compare to year 1 of modeled fishpond ammonium N to the 2014-2015 fishpond water ammonium data
fp_amm_y1 <- fp_water_df %>% filter(nitrogen == "Na_fp") %>% 
  filter(time >= 0 & time < 1.15)

ggplot(data = fp_amm_y1, aes(x = time, y = concentration)) +geom_line(col = 'red')

amm_combo_y1 = ggplot() + 
  geom_line(data = fp_amm_y1, aes(x = time*365, y = fp_amm_y1$concentration), col = 'red') + 
  geom_point(data = ammonium_data, aes(x = day_ct, y = amm_obs),col='magenta') +
  geom_errorbar(data = ammonium_data, aes(x = day_ct, ymin=amm_obs-amm_sd, ymax=amm_obs+amm_sd), width=3, col='magenta') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  guides(fill=guide_legend()) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

amm_combo_y1
```


```{r}
# compare to year 80 of modeled fishpond ammonium N to the 2014-2015 fishpond water ammonium data
fp_amm_y20 <- fp_water_df %>% filter(nitrogen == "Na_fp") %>% 
  filter(time >= 19 & time < 20.15)


amm_combo_y20 = ggplot() +
  geom_line(data = fp_amm_y20, aes(x = (time-19)*365, y = fp_amm_y20$concentration), col = 'red') + 
  geom_point(data = ammonium_data, aes(x = day_ct, y = amm_obs),col='magenta') +
  geom_errorbar(data = ammonium_data, aes(x = day_ct, ymin=amm_obs-amm_sd, ymax=amm_obs+amm_sd), width=3, col='magenta') +
  labs(x = 'Time (Days)', y = 'Concentration (mg/L)') +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5))

amm_combo_y20
```

Prepare the data to do the MSE analysis
```{r}
# need to average 4 times steps to get the average for that day. Assign a day number in integer form that can be compared against the days where there is data. Then make calculate MSE or some other metric for comparison

# create daily averages for nitrate concentrations modeled in the fishpond water
fp_nit_day <- data.frame(day_ct = seq(1,365*years),
                         nit_pred = 0)

for (i in 1:length(fp_nit_day$day)){
  fp_nit_day$nit_pred[i] =  (fp_nitrate6$concentration[i]+fp_nitrate6$concentration[i+1]+ fp_nitrate6$concentration[i+2]+fp_nitrate6$concentration[i+3])/4
}

# create daily averages for ammonium concentrations modeled in the fishpond water
fp_amm_day <- data.frame(day_ct = seq(1,365*years),
                         amm_pred = 0)

for (i in 1:length(fp_amm_day$day)){
  fp_amm_day$amm_pred[i] =  (fp_ammonium6$concentration[i]+fp_ammonium6$concentration[i+1]+ fp_ammonium6$concentration[i+2]+fp_ammonium6$concentration[i+3])/4
}

# for days outside of the data range of the first year, need to do some kind of offset (subtract 365*# of years different from the first year)

```

```{r}
amm_mse_df <- merge(ammonium_data, fp_amm_day, by = "day_ct")
nit_mse_df <- merge(nitrate_data, fp_nit_day, by = "day_ct")

amm_mse = MSE(y_pred = amm_mse_df$amm_pred, y_true = amm_mse_df$amm_obs)
nit_mse = MSE(y_pred = nit_mse_df$nit_pred, y_true = nit_mse_df$nit_obs)

amm_mse
nit_mse

# Nash-Sutcliffe Efficiency (NSE) is considered teh normalized MSE
# NSE = 1 - ( sum( (obs - sim)^2 ) / sum( (obs - mean(obs))^2 )
#Nash-Sutcliffe efficiencies range from -Inf to 1. Essentially, the closer to 1, the more accurate the model is. 
#-) NSE = 1, corresponds to a perfect match of modelled to the observed data. 
#-) NSE = 0, indicates that the model predictions are as accurate as the mean of the observed data, 
#-) -Inf < NSE < 0, indicates that the observed mean is better predictor than the model.

amm_var_mean = mean(amm_mse_df$amm_var)
nit_var_mean = mean(nit_mse_df$nit_var)

amm_nse = 1 - (amm_mse/amm_var_mean)
nit_nse = 1 - (nit_mse/nit_var_mean)

amm_nse # problematic that this is negative
nit_nse # probelmatic that this isn't that close to 1

# NSE seems to be used extensively in hydrologic models

# just by removing the outlier in the ammonium data set, the NSE went from  -1.138588 to 0.575
```

Total food production from the He‘eia IAAS

Agriculture production

```{r}
current_taro = 7300 # current area (m^2) in taro production

banana_area = 50181 # area (m^2) producing banana in 20 years
breadfruit_area = 50181 # area (m^2) producing breadfruit in 20 years
taro_area = 125453 # area (m^2) producing taro in 20 years

banana_yield = 1.49 # mean of 0.58-2.39 kg/yr from Bremer, et al (2018)
breadfruit_yield = 1.32 # mean of 0.20-2.43 kg/yr from Bremer, et al (2018)
taro_yield = 1.46 # mean of 1.12-1.79 kg/yr from Bremer, et al (2018)

taro_current_mass = current_taro*taro_yield

banana_mass = banana_area*banana_yield
breadfruit_mass = breadfruit_area*breadfruit_yield
taro_restored_mass = taro_area*taro_yield

```

Examine 10 year intervals for all land use scenarios. In the case of the restored scenario, the 10 years begin after 20 years of restoration are completed. Similarly for the urban land use scenario, the nitrogen loading used throughout the evaluated 10 year interval is constant, representing the loading from complete urbanization of the present.

```{r}
# extract the last value for the fish density in each land use scenario and multiply by He‘eia Fishpond's volume to get the mass in grams and convert to metric tons
current_fish <- df_control %>% 
  dplyr::select(time, catch) %>% 
  filter(time > 9) %>% 
  colSums() %>% 
  as.data.frame()

fish_ctrl = current_fish$.[2]*A_fp/10^6
 # check if it should be V_fp or A_fp at the end
restored_fish <- df_restored %>% 
  dplyr::select(time, catch) %>% 
  filter(time > 9) %>% 
  colSums() %>% 
  as.data.frame()

fish_rst = restored_fish$.[2]*A_fp/10^6

urban_fish <- df_urban %>% 
  dplyr::select(time, catch) %>% 
  filter(time > 9) %>% 
  colSums() %>% 
  as.data.frame()

fish_urb = urban_fish$.[2]*A_fp/10^6
```

```{r}

# create a dataset of the different crop data
Crop=c(rep("Taro", 3) , rep("Banana", 3) , rep("Breadfruit", 3) , rep("Fish", 3) )
Scenario =rep(c("Current" , "Restored" , "Urban") , 4)
Mass=c(taro_current_mass/1000, taro_restored_mass/1000, 0, 0, banana_mass/1000, 0, 0, breadfruit_mass/1000, 0, fish_ctrl, fish_rst, fish_urb)
crop_data=data.frame(Crop,Scenario,Mass)

# Stacked
ggplot(crop_data, aes(fill = Crop, y=Mass, x=Scenario)) + 
    geom_bar(stat="identity") +
    #scale_fill_discrete_qualitative(palette = "Cold") +
    guides(fill=guide_legend()) +
    labs(y = "Mass (Metric Tons)", x = "Land Use Scenario") +
    theme(axis.text.x= element_text(size = 8, hjust = .2), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5), legend.title.align = 0.5)
```

What if you were instead going to do a time series of 10 or 20 years from current conditions to either restored or urban conditions and you add a linear proportion of what the loading is supposed to increase by with each successive year?
