---
title: "vbgf_test"
author: "Nakoa Farrant"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(FSA)
library(FSAdata)
library(nlstools)
library(TropFishR)
library(tidyverse)
library(ggplot2)
library(plyr)
```






Found in Cheung, et al (2013) Nature Climate Change "Shrinking of fishes exacerbates impacts of global ocean changes on marine ecosystems". All parameters were determined from fish base. See supplementary materials from this paper

asymptotic weight relationship for each fish is given as W_infi
W_infi = a*(L_infi)^(b)

for Chanos chanos (milkfish), 
asymptotic length (L_infi) 183.4 cm, 
K = 0.10 1/year, 
a = 0.00193, 
b = 3.0 
where a and b are parameters of length-weight relationships for this fish. 

for Mugil cephalus (mullet)
asymptotic length (L_infi) 100.0 cm, 
K = 0.27 1/year, 
a = 0.0321 for M. cephalus found in Hawai'i. for overall M. cephalus in fishbase, a = 0.0085 
b = 3.0 

```{r}
t <- seq(0,60,0.1) # time step of 0.1 years for a 60 year time frame

Linf_chanos = 183.4 # cm
K_chanos = 0.10 # 1/year
a_chanos = 0.00193 
b_chanos = 3.0 
Winf_chanos = a_chanos*Linf_chanos^b_chanos # grams

Linf_mugil = 60 # cm, was 100cm, could also be 63 which is the mean of fishbase.60 top length of adults in Hawai'i was around 50cm but could bumped up to 
K_mugil = 0.27 # 1/year and could be 0.24 which is the geometric mean of fishbase
a_mugil = 0.0321 # was 0.0085 for overall fishbase sample. 
b_mugil = 3.0 
Winf_mugil = a_mugil*Linf_mugil^b_mugil # grams
# when using Linf = 100cm and a_mugil= 0.0085, Winf = 8500
# reported lengths from hawai'i estuary sample go to 51.6cm and weights of 1800 g
# using Linf 50cm and a = 0.0321, b = 3.0, Winf = 4012g which seems much higher than the sample

Wt_chanos <- VBGF(list(Winf=Winf_chanos, K = K_chanos), t = t) 
#plot(t, Wt_chanos, t="l", main = "VBGF for Chanos chanos (Milkfish)")

Wt_mugil <- VBGF(list(Winf=Winf_mugil, K = K_mugil), t = t) 
plot(t, Wt_mugil, t="l", main = "VBGF for Mugil cephalus (Mullet)")

# Use empirical method to estimate the natural mortality rates for each fish type
# This method is documented here https://cran.r-project.org/web/packages/TropFishR/vignettes/tutorial.html with reference to Then, et al 2015
# this is the TropFishR library

M_chanos = M_empirical(Linf = Linf_chanos, K_l = K_chanos, method = "Then_growth")
M_mugil = M_empirical(Linf = Linf_mugil, K_l = K_mugil, method = "Then_growth")
paste("M chanos =", as.numeric(M_chanos))
paste("M mugil =", as.numeric(M_mugil))

# Using the Pauly 1979 empirical equation for t0 from fishbase. Seems to be log10 and not ln because other instances on the fishbase page explicityly use ln. http://www.fishbase.us/manual/English/Key%20Facts.htm#resilience
t0_mugil = -exp(-0.3922-0.2752*log10(Linf_mugil)-1.038*log10(K_mugil))
Lm_mugil = 35 #cm, length at maturity

#calculate the time to maturity from empirical formula on fishbase 
tm_mugil = t0_mugil- log(1-Lm_mugil/Linf_mugil)/K_mugil

# length classwith the highest biomass in the unfished population
Lopt_mugil = Linf_mugil*(3/(3+M_mugil/K_mugil))
```

```{r}
mugil_list <- list(Winf = Winf_mugil, K = K_mugil, t0 = t0_mugil, M = M_mugil, tr = 2)

mugil_pred <- predict_mod(mugil_list, FM_change = seq(0.3,0.4,0.1), tc_change = seq(2,4,1), type = 'ypr')  #where it is maximal  = MSY

plot(mugil_pred, xaxis1 = "FM", yaxis_iso = "Lc", yaxis1 = "Y_R", yaxis2 = "B_R", mark = TRUE, ylab1 = "Y/R", ylab2 = "B/R")

# in the below graph with Y_R and B_R on yaxis 1 and 2, respectively, the biomass is just Y/R divided by fishing mortality
```


```{r}
data(hake)
hake_vpa = VPA(hake, terminalE = 0.5, analysis_type = "VPA", plot = TRUE,
catch_unit = "'000", plus_group = TRUE)
```


```{r}
Wt_mugil_df <- as.data.frame(t) %>% 
  mutate(weight = Wt_mugil, gr_rate = 0)

# calculate growth rate of fish in grams/6hr
for (i in 1:length(Wt_mugil_df$gr_rate)-1){
  Wt_mugil_df$gr_rate[i] = (Wt_mugil_df$weight[i+1]-Wt_mugil_df$weight[i])/0.1/365/4 
}

# calculate average growth rate during the first 4 years according to the von Bertalanffy
avg_gr_4 = mean(Wt_mugil_df$gr_rate[1:40])
paste(round(avg_gr_4, 3)) # [g/6hr]
```



```{r}
# using a different library but the same empirical method (Then growth) as above, find the same results for natural mortality rates
# https://rdrr.io/cran/fishmethods/man/M.empirical.html

# does the same mortality stuff from above

library(fishmethods)
Mchanos = M.empirical(Linf = Linf_chanos, Kl = K_chanos, method = 11)
Mmugil = M.empirical(Linf = Linf_mugil, Kl = K_mugil, method = 11)

paste("M.chanos =", as.numeric(Mchanos))
paste("M.mugil =", as.numeric(Mmugil))

```

```{r}
data(whiting)
View(whiting)

```

```{r}
# Virtual Popuation Analysis with age-composition data
output <- VPA(param = whiting, catch_columns = 1, terminalE = 0.5, analysis_type = "VPA")
plot(output)
```

```{r}
# Virtual population analysis with length-composition data
data(hake)
VPA(hake, terminalE = 0.5, analysis_type = "VPA", plot = TRUE,
    catch_unit = "'000", plus_group = TRUE)
```


```{r}
midlengths_chanos <- c(9, 15, 21, 27, 33, 39, 45, 51, 57, 63, 69, 75, 100, 160)
catch_chanos <- c(0, 0, 10000, 5000, 50000, 50000, 50000, 68990, 0, 0, 0, 0, 25000, 16000)
chanos_pred_list <- list(midLengths = midlengths_chanos, catch = catch_chanos, Linf = Linf_chanos, K = K_chanos, t0 = 0, M = M_chanos, FM = 0, a = a_chanos, b = b_chanos, meanValue = c(1.0, 1.0, 1.0, 1.5, 1.5, 2.0, 2.0, 2.5, 2.5, 2.5, 3.0, 3.0, 3.0, 3.0))
```


```{r}
TB1 <- VPA(chanos_pred_list, terminalE = 0.5, analysis_type = "VPA", plot = TRUE,
    catch_unit = "'000", plus_group = TRUE)
```


```{r, data frame organization}
fish_df = as.data.frame(t) %>% 
  # convert weights from grams to kg
  mutate("Chanos chanos" = Wt_chanos/1000, "Mugil cephalus" = Wt_mugil/1000) %>% 
  gather(key = Fish, value = weight, c("Chanos chanos", "Mugil cephalus"))
```

```{r, nice VBGF graph}


vbgf_plot <- ggplot(data = fish_df, aes(x = t, y = weight, col = Fish)) +
  geom_line() +
  labs(title = "von Bertalanffy Growth Models", x = "Age (years)", y = "Weight (kg)") +
  guides(fill = FALSE) +
  theme(axis.text.x= element_text(size = 8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(hjust = 0.5)) +
  scale_colour_hue(name="Fish Type",
                         breaks=c("Chanos chanos", "Mugil cephalus"),
                        labels = c(expression(italic("Chanos chanos")), 
        expression(italic("Mugil cephalus"))))

vbgf_plot
# somthing is wronge with this graph. It is either in kg or some other units. Also not sure about the x-axis
```



Searched the sea around us website 
for mugil cephalus
- maximum length is 67cm, trophic level is 2.26, medium demersals (30-89cm) functional group



```{r}
library(rfishbase)
```


```{r}
mugil_popgrowth = popgrowth(species_list = "Mugil cephalus")
```

```{r}
length_freq(species_list = "Mugil cephalus")
```

```{r}
stocks(species_list = "Mugil cephalus")
```

```{r}
popqb(species_list = "Mugil cephalus")
```

```{r}
maturity("Mugil cephalus")
```

Whole-body nitrogen and crude protein: L.J. Ramseyer (2000, in review; see also www.mi.nmfs.gov/Nfish.html) has analysed the relationship between whole-fish wet weight and whole-body nitrogen content for 68 species and hybrids, based on data extracted from the literature. He found the following relationship: log N (g) = 1.03*(log wet weight) - 1.65; n=2811, r^2=0.996, p<0.001. For the conversion from nitrogen to crude protein he gives the ratio: crude protein = 6.25 * nitrogen. We have added these relationships here for your convenience.


